### 2020-04-13 introduction.tex
minor grammatical edits

Since the Industrial Revolution in the 1800s, rapidly rising greenhouse gas
emissions have contributed significantly to global warming, the consequences
of which have been acutely felt in recent years through increasingly frequent
and record breaking extreme weather events. Urgent measures are necessary to
limit CO$_2$ emissions and the negative impacts of global warming.

Since electricity generation from burning fossil fuels represents the
largest source of CO$_2$ emissions (38\% in 2018 \cite{iea_global_2019}),
replacing them with low-carbon
alternatives would effectively curb a large fraction of these emissions. Out
of all low-carbon power sources, nuclear power in general is best suited to
replace fossil fuel burning; it provides consistent base-load
power independent of weather and local geographical conditions. Solar and wind
power are dependent on favorable weather conditions while hydropower requires
relatively more area and is limited to geologically appropriate locations.
This argument provides a strong case for nuclear power being the most
appropriate and realistic alternative to burning fossil fuels for our current
and future electricity needs.

Large scale reactor deployments would be necessary to displace the
presently large share of energy production from fossil fuel power plants;
however, several obstacles stand in the way of mass reactor deployments. These
include perceived safety risks, sustainability concerns, nuclear proliferation
risks, and the ability to compete economically with other sources of energy.
The \gls{MSR} concept, one of six advanced reactor designs
selected by the \gls{GIF} \cite{gif_technology_2002} for continued research
and development, is a potential solution to the aforementioned issues.

The primary coolant in MSRs is a molten salt mixture,
with fissile and/or fertile material directly dissolved in the coolant.
MSRs possess an inherently robust safety feature in the
strongly negative fuel temperature reactivity coefficient from Doppler
broadening and thermal fuel expansion that greatly reduces
the risk of a reactor power excursion. Many designs can also incorporate the
thorium fuel cycle for improved sustainability arising from the use of
abundant natural thorium resources and reduced transuranic waste. The
resultant reduced radiotoxicity from transuranic waste also reduces costs
associated with long-term nuclear waste storage. In addition, the ability to
operate at near atmospheric pressures eliminates the need for a thick pressure
vessel and drives down construction costs, while online fuel reprocessing
reduces reactor downtime during reactor operation.

However, the liquid fuel form also brought about novel computational
challenges in simulating the transient behavior of \glspl{MSR}; the
neutronics and thermal-hydraulics are more tightly coupled due to the strong
temperature reactivity coefficient, and the additional advection coupling
terms. Furthermore, we have to account for the movement of \glspl{DNP} as
they are now generated directly within the primary coolant loop. Therefore,
the choice of coupling methods for each set of physics requires careful
consideration. Existing reactor system-level codes and modeling approaches
for conventional \gls{LWR} analysis contain simplifying assumptions in
multiphysics coupling and other areas that render the techniques unsuitable
for simulating \glspl{MSR}. Thus, making minor modifications to these codes
without changing the underlying algorithms is not the best approach for
\gls{MSR} safety analysis.

In the past two decades, researchers have developed several
new tools for simulating steady-state and transient behavior in \glspl{MSR}.
Many of the earlier efforts featured simplifications in simulating
thermal-hydraulics by solving 1-D Navier-Stokes equations or using
predetermined uniform velocity fields \cite{krepel_dyn3d-msr_2007}
\cite{kophazi_development_2009}. In more recent years, there has been
significant progress towards fully coupled, spatial codes that feature
2-D axisymmetric or full 3-D models. In 2011, Cammi et al.
\cite{cammi_multi-physics_2011} performed a ``\gls{MPM}'' analysis
of a simplified 2-D axisymmetric model of a single \gls{MSBR} fuel channel
using the commercial finite element analysis software COMSOL Multiphysics. The
physics were implemented through the two-group neutron diffusion
equations, and the \gls{RANS} standard $k-\epsilon$ turbulence model, for the
neutronics and thermal-hydraulics respectively. The
authors emphasized the need for proper full coupling of the multiphysics and
presented both steady-state and transient results in various
scenarios such as reactivity insertions, changes in pumping rate, and the
presence of periodic perturbations. This approach was featured again in a
later paper by Fiorina et al. \cite{fiorina_modelling_2014} in 2014 for a 2-D
axisymmetric model of the
\gls{MSFR}. The authors presented results from the Politecnico di Milano
COMSOL-based approach, and another approach by researchers from Delft
University of Technology, where they coupled their in-house neutronics and
thermal-hydraulics codes, DALTON and HEAT respectively. With multigroup
neutron diffusion and \gls{RANS} formulations on ultra fine meshes, both
models showed good agreement in the steady-state neutron flux, temperature,
and \gls{DNP} distributions, and in the power responses following various
accident transient initiations. Aufiero et al. \cite{aufiero_development_2014}
concurrently developed
a full-core 3-D model of the \gls{MSFR} on OpenFOAM, albeit with one-group
neutron diffusion to reduce computational load. With the 3-D model, the
authors could simulate the asymmetric reactor response to the failure of a
single pump in the sixteen-pump \gls{MSFR} configuration. The authors also
provided some quantitative data supporting the use of implicit coupling over
explicit coupling to obtain accurate solutions of the transient cases.
Recognizing the huge computational burden required for full 3-D simulations, 
later authors came up with innovative ways to alleviate this issue such as
selective geometrical reduced order modeling for various components of a
reactor based on the importance of the physical phenomena being simulated
\cite{zanetti_geometric_2015}, or using a novel, efficient method for
neutronics calculations \cite{laureau_transient_2017}.

>>>>

Since the Industrial Revolution in the 1800s, rapidly rising greenhouse gas
emissions have contributed significantly to global warming. The consequences
have been acutely felt in recent years through increasingly frequent
and record breaking extreme weather events. Urgent measures are necessary to
limit CO$_2$ emissions and the negative impacts of global warming.

Since electricity generation from burning fossil fuels represents the
largest source of CO$_2$ emissions (38\% in 2018 \cite{iea_global_2019}),
replacing them with low-carbon
alternatives would effectively curb a large fraction of emissions. Out
of all low-carbon power sources, nuclear power is arguably best suited to
replace fossil fuel burning; it provides consistent base-load
power independent of weather and local geographical conditions. Solar and wind
power are dependent on favorable weather conditions while hydropower requires
relatively more area and is limited to geologically appropriate locations.
This argument provides a strong case for nuclear power being the most
appropriate and realistic alternative to burning fossil fuels for our current
and future electricity needs.

Large-scale reactor deployments are necessary to displace the
presently large share of energy production from fossil fuel power plants.
However, several obstacles stand in the way of mass reactor deployments. These
include perceived safety risks, sustainability concerns, nuclear proliferation
risks, and the ability to compete economically with other sources of energy.
The \gls{MSR} concept, one of six advanced reactor designs
selected by the \gls{GIF} \cite{gif_technology_2002} for continued research
and development, is a potential solution to the aforementioned issues.

The primary coolant in MSRs is a molten salt mixture
with fissile and/or fertile material directly dissolved in the coolant.
MSRs possess an inherently robust safety feature in the
strongly negative fuel temperature reactivity coefficient from Doppler
broadening and thermal fuel expansion that greatly reduces
the risk of a reactor power excursion. Many designs can also incorporate the
thorium fuel cycle for improved sustainability arising from the use of
abundant natural thorium resources and reduced transuranic waste. The
resultant reduced radiotoxicity from transuranic waste also reduces costs
associated with long-term nuclear waste storage. In addition, the ability to
operate at near atmospheric pressures eliminates the need for a thick pressure
vessel and drives down construction costs, while online fuel reprocessing
reduces reactor downtime during reactor operation.

However, the liquid fuel form also brings about novel computational
challenges in simulating the transient behavior of \glspl{MSR}; the
neutronics and thermal-hydraulics are more tightly coupled due to the strong
temperature reactivity coefficient, and the additional advection coupling
terms. Furthermore, we have to account for the movement of \glspl{DNP} as
they are now generated directly within the primary coolant loop. Therefore,
the choice of coupling methods for each set of physics requires careful
consideration. Existing reactor system-level codes and modeling approaches
for conventional \gls{LWR} analysis contain reactor-specific assumptions in
multiphysics coupling and other areas that render the techniques less suitable
for simulating \glspl{MSR}. Thus, making minor modifications to these codes
without changing the underlying algorithms is not the best approach for
\gls{MSR} safety analysis.

In the past two decades, researchers have developed several
new tools for simulating steady-state and transient behavior in \glspl{MSR}.
Many of the earlier efforts featured simplifications in simulating
thermal-hydraulics by solving 1-D Navier-Stokes equations or using
predetermined uniform velocity fields \cite{krepel_dyn3d-msr_2007}
\cite{kophazi_development_2009}. In more recent years, there has been
significant progress towards fully coupled, spatial codes that feature
2-D axisymmetric or full 3-D models. In 2011, Cammi et al.
\cite{cammi_multi-physics_2011} performed a ``\gls{MPM}'' analysis
of a simplified 2-D axisymmetric model of a single \gls{MSBR} fuel channel
using the commercial finite element analysis software COMSOL Multiphysics. The
physics were implemented through the two-group neutron diffusion
equations, and the \gls{RANS} standard $k-\epsilon$ turbulence model, for the
neutronics and thermal-hydraulics respectively. The
authors emphasized the need for proper full coupling of the multiphysics and
presented both steady-state and transient results in various
scenarios such as reactivity insertions, changes in pumping rate, and the
presence of periodic perturbations. This approach was featured again in a
later paper by Fiorina et al. \cite{fiorina_modelling_2014} in 2014 for a 2-D
axisymmetric model of the
\gls{MSFR}. The authors presented results from the Politecnico di Milano
COMSOL-based approach, and another approach by researchers from Delft
University of Technology, in which they coupled their in-house neutronics and
thermal-hydraulics codes, DALTON and HEAT respectively. With multigroup
neutron diffusion and \gls{RANS} formulations on ultra fine meshes, both
models showed good agreement in the steady-state neutron flux, temperature,
and \gls{DNP} distributions, and in the power responses following various
accident transient initiations. Aufiero et al. \cite{aufiero_development_2014}
concurrently developed
a full-core 3-D model of the \gls{MSFR} on OpenFOAM, albeit with one-group
neutron diffusion to reduce computational load. With the 3-D model, the
authors could simulate the asymmetric reactor response to the failure of a
single pump in the sixteen-pump \gls{MSFR} configuration. The authors also
provided some quantitative data supporting the use of implicit coupling over
explicit coupling to obtain accurate solutions of the transient cases.
Recognizing the huge computational burden required for full 3-D simulations, 
later authors came up with innovative ways to alleviate this issue such as
selective geometrical reduced order modeling for various components of a
reactor based on the importance of the physical phenomena being simulated
\cite{zanetti_geometric_2015}, or using a novel, efficient method for
neutronics calculations \cite{laureau_transient_2017}.

### 2020-04-13 msr.tex
Completed first draft of Chapters 1-4 for review
minor grammatical edits

The first \gls{MSR}, named the \gls{ARE}, dates back to the 1940s,
as part of the US Aircraft Nuclear Propulsion program
\cite{rosenthal_molten-salt_1970}; the molten salt concept was considered due
to the stability of molten salts at high temperatures and neutron radiation.
The 2.5 MW$_{\text{th}}$ reactor was built at \gls{ORNL}, where it achieved
criticality on November 1954 and generated 100 MWh over nine days
\cite{rosenthal_molten-salt_1970}. The fuel
consisted of enriched uranium in a molten salt mixture of NaF, ZrF$_4$, and
UF$_4$, and was moderated by blocks of beryllium oxide. The project ultimately
never came to fruition as the development of intercontinental ballistic
missiles effectively eliminated the need for long-range nuclear-powered
bomber aircraft.

@@

Following a lull lasting through the late 20th century, \gls{MSR} research
picked up pace due to renewed interest initiated by the \gls{GIF} in 2001.
Today, there are numerous \gls{MSR} concepts under active development led by
various national and commercial bodies. Many differences exist between
different concepts and the only common denominator is the use of nuclear fuel
dissolved in molten salt. Table * is a non-exhaustive list of \glspl{MSR}
under development.

@@

The flexibility of \glspl{MSR} is best illustrated by the various designs
under development today. Graphite-moderated thermal-spectrum \glspl{MSR} are
typically straightforward \gls{LEU} burners, or $^{232}$Th/$^{233}$U
iso-breeders/breeders, while epithermal- and fast-spectrum \glspl{MSR} have
the additional options of operating as \gls{TRU} fuel burners or
$^{238}$U/$^{239}$Pu breeders. Breeder designs can be further categorized into
one- or two-fluid designs; two-fluid designs feature separate blanket molten
salt mixtures that contain higher proportions of fertile material than the
fuel salt mixture.

@@

\glspl{MSR} may be more resilient to pump failure accidents as natural
circulation could passively sustain enough heat transfer through the heat
exchanger to remove any remaining fission and decay heat after the failure. If
natural circulation proves insufficient, the aforementioned freeze plug can
drain the salt out of the core. Decay heat in \glspl{MSR} with online
reprocessing is typically lower than that in \glspl{LWR} due to the
continuous removal of \glspl{FP}. For example, the decay heat in an \gls{MSFR}
after reaching equilibrium salt composition is approximately 3.5\% compared
to 6\% in \glspl{LWR} \cite{brovchenko_design-related_2013}.

>>>>

The first \gls{MSR}, named the \gls{ARE}, dates back to the 1940s
as part of the US Aircraft Nuclear Propulsion program
\cite{rosenthal_molten-salt_1970}; the molten salt concept was considered due
to the stability of molten salts at high temperatures and neutron radiation.
The 2.5 MW$_{\text{th}}$ reactor was built at \gls{ORNL}, where it achieved
criticality on November 1954 and generated 100 MWh over nine days
\cite{rosenthal_molten-salt_1970}. The fuel
consisted of enriched uranium in a molten salt mixture of NaF, ZrF$_4$, and
UF$_4$, and was moderated by blocks of beryllium oxide. The project ultimately
never came to fruition as the development of intercontinental ballistic
missiles effectively eliminated the need for long-range nuclear-powered
bomber aircraft.

@@

Following a lull lasting through the late 20th century, \gls{MSR} research
picked up pace due to renewed interest initiated by the \gls{GIF} in 2001.
Today, there are numerous \gls{MSR} concepts under active development led by
various national and commercial bodies. Many differences exist between
different concepts and the only common denominator is the use of nuclear fuel
dissolved in molten salt.
%Table * is a non-exhaustive list of \glspl{MSR} under development.

@@

The flexibility of \glspl{MSR} is best illustrated by the various designs
under development today. Graphite-moderated thermal-spectrum \glspl{MSR} are
typically straightforward \gls{LEU} burners or $^{232}$Th/$^{233}$U
iso-breeders/breeders, while epithermal- and fast-spectrum \glspl{MSR} have
the additional options of operating as \gls{TRU} fuel burners or
$^{238}$U/$^{239}$Pu breeders. Breeder designs can be further categorized into
one- or two-fluid designs; two-fluid designs feature separate blanket molten
salt mixtures that contain higher proportions of fertile material than the
fuel salt mixture.

@@

\glspl{MSR} may be more resilient to pump failure accidents as natural
circulation could passively sustain enough heat transfer through the heat
exchanger to remove any remaining fission and decay heat after the failure. If
natural circulation proves insufficient, the aforementioned freeze plug can
drain the salt out of the core. Decay heat in \glspl{MSR} with online
reprocessing is typically lower than that in \glspl{LWR} due to the
continuous removal of \glspl{FP}. For example, the decay heat in an \gls{MSFR}
after reaching equilibrium salt composition is expected to be approximately
3.5\% of full reactor power compared to 6\% in \glspl{LWR}
\cite{brovchenko_design-related_2013}.
